Consistent Polyhedral Surrogates via Embedding

### 2021-07-28 colt-21/rebuttal.tex

\documentclass[12pt]{colt2021} %% Anonymized submission
      \usepackage{times}
      \usepackage{lmodern}
      \usepackage{hyperref}       % hyperlinks  %[implicit=false, bookmarks=false]
      \usepackage{booktabs}       % professional-quality tables
      \usepackage{amsfonts}       % blackboard math symbols
      \usepackage{nicefrac}       % compact symbols for 1/2, etc.
      \usepackage{microtype}      % microtypography

      \PassOptionsToPackage{numbers, sort, compress}{natbib}

      \usepackage{mathtools, verbatim}
      %\usepackage[thmmarks, thref, amsthm]{ntheorem}
      \usepackage{color}
      \definecolor{darkblue}{rgb}{0.0,0.0,0.2}
      \hypersetup{colorlinks,breaklinks,
      	linkcolor=darkblue,urlcolor=darkblue,
      	anchorcolor=darkblue,citecolor=darkblue}
      \usepackage{wrapfig}
      \usepackage[small]{caption}
      % NOTE: had to comment out "todo" in colt2021.cls
      \usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
      %\usepackage{accents}
      \usepackage{bbm}
      \usepackage{xspace}
      \usetikzlibrary{arrows}
      \usetikzlibrary{shapes}

      \makeatletter
      \let\Ginclude@graphics\@org@Ginclude@graphics 
      \makeatother

      %\usetikzlibrary{calc}
      \newcommand{\Comments}{0}
      \newcommand{\mynote}[2]{\ifnum\Comments=1\textcolor{#1}{#2}\fi}
      \newcommand{\mytodo}[2]{\ifnum\Comments=1%
      	\todo[linecolor=#1!80!black,backgroundcolor=#1,bordercolor=#1!80!black]{#2}\fi}
      \newcommand{\raf}[1]{\mynote{green}{[RF: #1]}}
      \newcommand{\raft}[1]{\mytodo{green!20!white}{RF: #1}}
      \newcommand{\jessie}[1]{\mynote{purple}{[JF: #1]}}
      \newcommand{\jessiet}[1]{\mytodo{purple!20!white}{JF: #1}}
      \newcommand{\bo}[1]{\mynote{blue}{[Bo: #1]}}
      \newcommand{\botodo}[1]{\mytodo{blue!20!white}{[Bo: #1]}}
      \newcommand{\btw}[1]{}%\mytodo{orange!80!white}{BTW: #1}}
      \newcommand{\discuss}[1]{\mynote{cyan!20!white}{#1}}
      \ifnum\Comments=1               % fix margins for todonotes
      \setlength{\marginparwidth}{1in}
      \fi

      \newcommand{\reals}{\mathbb{R}}
      \newcommand{\posreals}{\reals_{>0}}%{\reals_{++}}
      \newcommand{\simplex}{\Delta_\Y}
      \newcommand{\relint}[1]{\mathrm{relint}(#1)}
      \newcommand{\interior}{\mathrm{int}\,}
      \newcommand{\prop}[2][\mathcal{P}]{\mathrm{prop}_{#1}[#2]}
      \newcommand{\elic}{\mathrm{elic}}
      \newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
      \newcommand{\conscvx}{\mathrm{cons}_\mathrm{cvx}}
      \newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
      \newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}
      \newcommand{\ccdim}{\mathrm{cc\,dim}}
      \newcommand{\rank}{\mathrm{rank}}
      \newcommand{\proj}{\mathrm{proj}}
      \newcommand{\supp}{\mathrm{supp}}
      \newcommand{\spn}{\mathrm{span}}

      \newcommand{\range}{\mathrm{range}\,}
      \newcommand{\zeros}[1]{\mathrm{ker}_\P\,#1}
      \newcommand{\codim}{\mathrm{codim}}
      \newcommand{\Pcodim}{\mathcal{P}\!\text{-}\mathrm{codim}}
      \newcommand{\Pcodimension}{$\mathcal{P}$-codimension\,}

      \newcommand{\propdis}{\mu}
      \newcommand{\affhull}{\mathrm{affhull}}
      \newcommand{\epi}{\mathrm{epi}}
      \newcommand{\cl}{\mathbf{cl}}
      %\newcommand{\span}{\mathrm{span}}

      \newcommand{\A}{\mathcal{A}}
      \newcommand{\C}{\mathcal{C}}
      \newcommand{\D}{\mathcal{D}}
      \newcommand{\E}{\mathbb{E}}
      \newcommand{\F}{\mathcal{F}}
      \renewcommand{\L}{\mathcal{L}}
      \newcommand{\Lcvx}{\mathcal{L}^{\mathrm{cvx}}}
      \newcommand{\I}{\mathcal{I}}
      \newcommand{\N}{\mathcal{N}}
      \newcommand{\R}{\mathcal{R}}
      \renewcommand{\P}{\mathcal{P}}
      \newcommand{\Sc}{\mathcal{S}}  % jessie, feel free to redef, just not \S :-)
      \newcommand{\Scr}{\mathcal{S}}  % using this for the span of F_p - not sure if we want this to be parallel to the feasible subspace notation
      \newcommand{\U}{\mathcal{U}}
      \newcommand{\V}{\mathcal{V}}
      \newcommand{\X}{\mathcal{X}}
      \newcommand{\Y}{\mathcal{Y}}

      \newcommand{\ellbar}{\underline{\ell}}
      \newcommand{\lbar}{\underline{L}} % couldn't do L* while proofreading...
      \newcommand{\iden}{\mathrm{iden}}
      \newcommand{\im}{\mathrm{im}}
      \newcommand{\Var}{\mathrm{Var}}
      \newcommand{\CVaR}{\mathrm{CVaR}}

      \newcommand{\exploss}[3]{\E_{#3} #1(#2,Y)}
      \newcommand{\risk}[1]{\underline{#1}}
      \newcommand{\Ind}[1]{\mathbf{I}\{{#1}\}}
      \newcommand{\inprod}[2]{\langle #1, #2 \rangle}
      \newcommand{\toto}{\rightrightarrows}
      \newcommand{\ones}{\mathbbm{1}}

      %\newtheorem{theorem}{Theorem}
      %\newtheorem{lemma}{Lemma}
      %\newtheorem{proposition}{Proposition}
      %\newtheorem{corollary}{Corollary}
      %\newtheorem{conjecture}{Conjecture}
      %\newtheorem{definition}{Definition}
      \newtheorem{assumption}{Assumption}
      \newtheorem{condition}{Condition}
      %\newtheorem{remark}{Remark}

      \DeclareMathOperator*{\argmax}{arg\,max}
      \DeclareMathOperator*{\argmin}{arg\,min}
      \DeclareMathOperator*{\arginf}{arg\,inf}
      \DeclareMathOperator*{\sgn}{sgn}

      \usepackage{thmtools, thm-restate}
      %\declaretheorem{corollary}

      \begin{document}
      \title{}                     %% This removes title for more space

We thank the reviewers for their thoughtful comments.  On novelty of our results: We provide novel contributions both to specific elicitation/consistency bounds of interest (in Quadrant 4, not Quadrant 1 where reviewers seemed to focus), and conceptually to the broader literature on consistency of surrogate losses.  We outline both of these contributions below, and would urge the reviewers to consider both.

\paragraph{Novel bounds: Bayes risks and other risk measures}  
Our paper resolves an open question posed by Frongillo-Kash (NeurIPS 2015), on developing bounds for elicitation complexity for continuous estimation problems (Quadrants 3 and 4) with respect to convex surrogates.  (Our other novel bounds are mainly to illustrate the power of the tools, as we discuss below.)  Specifically, we show the first bounds on elicitation complexity of risk measures with respect to convex losses.  While the tools we use to show these bounds are heavily inspired by ideas in prior works, they required original and careful adaptation at every level; see Section 6 and Appendix C.

\paragraph{Conceptual: broader, easier, and stronger tool}  While our techniques draw inspiration from Ramaswamy-Agarwal (2016) and Agarwal-Agarwal (2015), we provide a significant conceptual contribution.  Specifically, we give a general tool to reason about consistency, which is:  (1) \underline{Broader}: the above works' definitions and results are tailored to Quadrant 1 (discrete, given target loss), while our tool applies to all 4 quadrants.  (2) \underline{Easier}: AA15 ``lifts'' property elicitation up to calibration, whereas we ``pull'' calibration down to elicitation, which is simpler to work with, as one need only reason about exact minimizers of expected loss rather than approximate minimizers.  We believe researchers will find this approach significantly simpler to conceptualize and apply.  Yet surprisingly, it is: (3) \underline{Stronger}: our example lower bound for abstain loss is stronger than what can be obtained via previous tools, namely feasible subspace dimension (FSD).  This extra strength stems from the fact that FSD relies only on local information (in the probability simplex), whereas we look at global information.  The combination of (1,2,3) is exactly what allowed us to prove the novel bounds for risk measures above.

\vspace{-0.2cm}
\paragraph{Other points / questions raised:}
\begin{itemize}
	\item {\color{red}R2} The work on composite proper losses from Williamson and collaborators is focused on eliciting the entire label distribution, rather than a property of that distribution as in our setting.  However, in eliciting the entire distribution, we agree the concepts are very similar, and will acknowledge this line of work.

	\item {\color{red} R2} and {\color{blue} R3}  That calibration implies indirect elicitation follows immediately from the definition.  To see that indirect elicitation can be strictly weaker, consider hinge loss with the link $\psi(u) = 2 \Ind{u < 1} - 1$.  We have hinge loss indirectly eliciting the mode via this link, as the linked minimizer of the surrogate always maps to the most likely outcome.  However, this link is not calibrated, as the excess risk by target loss misprediction can approach 0 by predicting u arbitrarily close to 1 for any $p$ such that $Pr[Y=1] > 1/2$.

	\item {\color{red}R2} We agree, the concept of indirect elicitation dates back to Lambert et al. (EC 2008) and perhaps before, and connections to consistency were clarified in AA15.  Since our literature review is mixed with the background section, these references were delayed; we will include them earlier.

\end{itemize}

      \end{document}

### 2021-07-29 notes/top-k.tex
Updating whiteboard writeup to clarify and think I have a proof sketch for hat gamma; see Zulip

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and narrow down to a finite representative set $\R$ for $L^{top-k}$.
Consider $\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}$.

\begin{claim}
	For each $p \in \mathrm{relint}\simplex$, the vertices of the expected loss $\inprod{p}{L^{top-k}(u)} = 1 - \inprod{p}{u} + \sum_{i=1}^k u_{[i]} - \inprod{p}{\min(u,\ones)}$ are denoted by $\R$.
\end{claim}
\begin{proof}
	First, observe the loss is affine in the first two terms for all $u \in \U$, and non-linearity is only introduced in the last two terms.
	Without loss of generality, we will fix a partition $\pi$ of the elements of $u$, and let $T_k(u)$ be the top $k$ elements of $u$ according to the partition $\pi$.
	Since $u \in \U \supseteq R^{\tt low}$, we have $u_{\pi_{k+1}} = \ldots = u_{\pi_n} = 0$, and take $k$ constraints constructing $R^{\tt low}$ given by $0 \leq u_{\pi_i}$ for $i \in [k]$.
	Call $T$ the top $k$ elements of the partition $\pi$.

	Moreover, there are $k$ constraints constructing $R^{\tt high}$, given by $	\forall i \in T_k(u), \,\,
	u_{i} \leq 1 + \frac{1}{k-1}\sum_{j \in T_k(u) \setminus \{i\}} u_j$.

	%need to introduce the segments where L is affine.
	Within each permutation $\pi$, we know that the third term is affine on each set $\{u \in \U \mid $ the top $k$ elements of $u$ are equal to $T\}$.

	Finally, we can consider affineness in the last term, where there is a ``switch'' at $u_y = 1$ for each $y$.
	There are $2^k$ choices of sets $V \in \{A,B\}^k$, $u \in V$ if and only if, for all $y \in \Y$, $V_y = A$ means $u_y \leq 1$ and $V_y = B$ otherwise. 
	We have $L$ affine on the region $\{u \in \U \mid u \in V\}$ for each choice of $V \in \{A,B\}^k$.
	We will generate such sets using $T$ and $S$, the latter of which is introduced below.

	%Now, since each of these inequalities generates a halfspace in $\reals^k$, and a vertex in $\reals^k$ is the intersection of $k$ such halfspaces, we observe that there are $2^k$ possible vertices (though some of these are redundant).

	For a fixed $T$ (the top $k$ elements of $\pi$), iterate over every $S \subseteq T$.
	Suppose $S$ is the set of elements of $u$ such that $0 \leq u_y \leq 1$ for all $y \in S$, and additionally have the constraints $1 \leq u_y \leq 1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j$.
	For each $T,S$ combination, we then have vertices at each of these $2^k$ possible equalities.
	\begin{align*}
	\forall i \in S, \,\, 0 &\leq u_i \leq 1 \\
	\forall i \in T \setminus S, \,\, 1 &\leq u_i \leq 1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j 
	\end{align*}
	Iterating over each of these $2^k$ inequalities, we see that vertices are generated at points $0,1$ or some constant $c_{TS}$ for each choice of inequalities for $T$ and $S$.

	Now, for each $T$ and $S$ and choice of equalities in each, let $M = \{y \mid u_y = 1\}$, and $H = \{y \in \Y \mid u_y = 1 + \sum_{j \in T_k(u) \setminus \{i\}} u_j\}$.
	(Note that $y \in H \implies u_y > 1$ unless $M$ is empty \jessie{and $H = \{y\}$?}, in which case, we will say $y \in M$ and $H = \emptyset$.) 
	\jessie{Be careful about if $c_{MH} = 1$}
	By construction of $T$ and $S$, we satisfy the set requirements of $\R$, and observe that for all $y \in H$, we have $u_y = 1 + \frac 1 {k-1}\sum_{j \in T_k(u) \setminus \{y\} u_j} = \frac{|M| + k -1}{k - |H|}$.

	As there are only finitely many choices of $T$ and $S$, a finite number of vertices generated by each of those choices, and they generate the sets generating $\R$.
	Moreover, for each $H,M$ generating an element of $R$, there is a $T,S$ generating the element of $\R$.
	Thus, we conclude $\R$ is the finite set of vertices of $\inprod{p}{L(\cdot)}$ for distributions $p \in \mathrm{relint}(\simplex)$.
\end{proof}

@@

\begin{claim}
	Fix $p \in \simplex$.
	$\hat \Gamma(p) \in \U$.
\end{claim}	
\begin{proof}
	Fix $p \in \simplex$, and take $H := |H_p|, M:= |M_p|$, and $B := |B_p|$.
	It suffices to show $H + M + B \leq k$ to show $\hat \Gamma (p) \in \R^{\tt low}$.

@@

The finite loss $L|_{\R}$ embedded by $L^{top-k}$ by
\begin{align}
L|_{\R}(r,y) &= \begin{cases}
0 & r_y = \bar r_{-y} + 1\\
\bar r - \frac 1 k & r_y = 1\\
1 + \bar r & r_y = 0
\end{cases}~.~
\end{align}

@@

>>>>

We will start with $\U$ as we show it is representative in Lemma~\ref{lem:top-threshold}, and the set of vertices $\R$ for $L^{top-k}$.
Consider $\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}$.

\begin{claim}
	For each $p \in \mathrm{relint}\simplex$, the vertices of the expected loss $\inprod{p}{L^{top-k}(u)} = 1 - \inprod{p}{u} + \sum_{i=1}^k u_{[i]} - \inprod{p}{\min(u,\ones)}$ are denoted by $\R$.
\end{claim}
\begin{proof}
	First, observe the loss is affine in the first two terms for all $u \in \U$, and non-linearity is only introduced in the last two terms.
	Fixing a set $T$ of size $k$, and for any $u \in \U$, we will call $T_k(u)$ the top $k$ elements of $u$.
	Moreover, denote by $A^T = \{u \in \U \mid T_k(u) = T\}$ as the set of $u$ whose top $k$ elements are exactly the elements of $T$.
	By construction, we now have $\bar u$ is affine in $A^T$ for each $T$.

	Since $u \in \U \supseteq R^{\tt low}$, we have $u_{[k+1]} = \ldots = u_{[n]} = 0$, and we also have $k$ constraints constructing $R^{\tt low}$ given by $0 \leq u_{i}$ for $i \in T$.	
	Moreover, there are $k$ constraints constructing $R^{\tt high}$, given by $\forall i \in T, \,\,
	u_{i} \leq 1 + \frac{1}{k-1}\sum_{j \in T_k(u) \setminus \{i\}} u_j$.

	Finally, we can consider affineness in the last term, where there is a ``switch'' at $u_y = 1$ for each $y$.
	For a fixed set $T$ and $u \in A^T$, consider any $S \subseteq T$.
	We can now construct the region $A^{T,S} = \{u \in A^T \mid u_y \in [0,1] \forall y \in S \wedge u_y \in [1,1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j] \forall y \in T \setminus S \}$.
	%  where for any $y \in S$, we have $u_y \in [0,1]$, and for $y \in T \setminus S$, we have $u_y \in [1,1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j]$.
	%There are $2^k$ choices of sets $V \in \{A,B\}^k$, $u \in V$ if and only if, for all $y \in \Y$, $V_y = A$ means $u_y \leq 1$ and $V_y = B$ otherwise. 
	As $A^{T,S}$ is a subset of $A^T$, we know the third term is affine on $A^{T,S}$, and construct $A^{T,S}$ so that it is affine on the fourth term: therefore, $L$ is affine on $A^{T,S}$ for all $T$ and $S \subseteq T$. 
	We now proceed to enumerate the vertices of $L$ by finding $k$ equalities imposed on $A^{T,S}$.
	(Note that technically we need $n$ hyperplanes to intersect to form a vertex, but with $T$ fixed, the other $n-k$ come from the requirement $u_i = 0 \forall i \in [n] \setminus T$.)

	%Now, since each of these inequalities generates a halfspace in $\reals^k$, and a vertex in $\reals^k$ is the intersection of $k$ such halfspaces, we observe that there are $2^k$ possible vertices (though some of these are redundant).
	For each $T,S$ combination, we then have vertices at each of these $2^k$ possible equalities, given by the following constraints.
	\begin{align*}
	\forall i \in S, \,\, 0 &\leq u_i \leq 1 \\
	\forall i \in T \setminus S, \,\, 1 &\leq u_i \leq 1 + \frac 1 {k-1} \sum_{j \in T_k(u) \setminus \{i\}} u_j 
	\end{align*}
	Iterating over each of these $2^k$ inequalities, we see that vertices are generated at points $0,1$ or some constant $c_{T,S}$ for each choice of inequalities for $T$ and $S$.

	Now, for fixed $T$, $S$, and choice of equalities in each, let $M = \{y \mid u_y = 1\}$, and $H = \{y \in \Y \mid u_y = 1 + \sum_{j \in T_k(u) \setminus \{i\}} u_j\}$.
	(Note that $y \in H \implies u_y > 1$ unless $M$ is empty \jessie{and $H = \{y\}$?}, in which case, we will say $y \in M$ and $H = \emptyset$.) 
	\jessie{Be careful about if $c_{MH} = 1$}
	By construction of $T$ and $S$, we satisfy the set requirements of $\R$, and observe that for all $y \in H$, we have $u_y = 1 + \frac 1 {k-1}\sum_{j \in T_k(u) \setminus \{y\} u_j} = \frac{|M| + k -1}{k - |H|}$.

	As there are only finitely many choices of $T$ and $S$, a finite number of vertices generated by each of those choices, and they generate the sets generating $\R$.
	Moreover, for each $H,M$ generating an element of $R$, there is a $T,S$ generating the element of $\R$.
	Thus, we conclude $\R$ is the finite set of vertices of $\inprod{p}{L(\cdot)}$ for distributions $p \in \mathrm{relint}(\simplex)$.
	%(If $p \not \in \mathrm{relint}(\simplex)$, then some points in $\R$ may not be vertices since any $T$ containing $y$ outside the support of $p$ will not be it dude )
\end{proof}

\begin{corollary}
	$L^{top-k}$ embeds $L^{top-k}|_\R$.
\end{corollary}

\begin{align}
L|_{\R}(r,y) &= \begin{cases}
0 & r_y = \bar r_{-y} + 1\\
\bar r - \frac 1 k & r_y = 1\\
1 + \bar r & r_y = 0
\end{cases}~.~
\end{align}

@@

\begin{claim}
	Fix $p \in \simplex$.
	$\hat \Gamma(p) \in \R$.
\end{claim}	
\begin{proof}
	Fix $p \in \simplex$, and take $H := |H_p|, M:= |M_p|$, and $B := |B_p|$.
	By construction, we just need to show that $H + M + B \leq k$ and we are done.

@@

@@

\section{One-pager}
Top-$k$ prediction is often used for image classification where labels may be ambiguous.
This generalization of 0-1 loss allows the algorithm to guess $k$ labels, and deeming a prediction ``correct'' if the true, assigned label is in the $k$-element set of predicted labels.
The top-$k$ discrete loss with predictions $r \in \R := \{R \subseteq \Y \mid |R| \leq k\}$ can be defined as follows:
\begin{align*}
\ell^k(r)_y &=
\begin{cases}
0 & y \in r \\
1 & \text{otherwise}
\end{cases}
\end{align*}

Like 0-1 loss, $\ell^k$ is $NP$-hard to optimize in most cases, and we want a convex surrogate for $\ell^k$ that is calibrated.
\citet{lapin2015top, lapin2015loss} propose a few natural convex surrogates for $\ell^k$, leaving calibration as an open question.
\citet{yang2020consistency} later show that their convex surrogates are not calibrated with respect to $\ell^k$ (summary of results in Table 1, functions $\psi_2, \psi_3$).
This begs the question: what are these surrogates calibrated with respect to?
We focus on one of the surrogates of \citet{lapin2015top,lapin2015loss}, denoted $L^{top-k}$ here, and find a finite representative set for it, giving a loss that $L^{top-k}$ embeds.
Note that for any $u \in \reals^n$, we denote $u_{[i]}$ to be the $i^{th}$ largest element of $u$.
\begin{align*}
L^{top-k}(u)_y &= \left(1 - u_y + \sum_{i=1}^k(u-e_y)_{[i]}\right)_+
\end{align*}

We enumerate the vertices of $L^{top-k}$ by first reducing to an infinite representative set $\U$ which allows us to simplify the form of $L^{top-k}$, then define the regions for which $\inprod{p}{L(\cdot)}$ is affine (for any $p$) so that we may enumerate the vertices of $L^{top-k}$.
These vertices have each coordinate ``low'' (e.g., $u_i = 0$), ``medium'' ($u_i = 1$), or ``high'' ($u_i$ is as high as possible while still being in $\U$).
Details are deferred to Appendix~\ref{app:top-k}.

\jessie{Bonus content: if $\hat \Gamma(p)$ pans out, then we have the non-redundancy result}

\jessie{Maybe: for intuition, keep example with $k = 2$ and $n = 3$?}
